\documentclass[11pt]{article}
\usepackage{a4}

\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\textwidth}{1.0in}
\addtolength{\headheight}{0.2in}

% So that the headers use the new dimensions, the package is
% included after the dimensions are reset:
\usepackage{fancyheadings}
\pagestyle{fancyplain}

\def\omniParTcl{{\tt omniParTcl}}
\def\tcltk{{\tt Tcl/Tk}}
\def\tcl{{\tt Tcl}}
\def\tk{{\tt Tk}}

\def\CXX{{\tt C++}}

\begin{document}
  
\title{{\bf \omniParTcl}\\
       {\normalsize Creating \tcltk\ GUIs for \CXX\ applications}}
\author{Kenneth R.\ Wood\\
        ORL\\
	{\small (ODE version 4.0 revisions by Tristan Richardson)}}
\date{\today} 
\maketitle

\begin{abstract}
\noindent
We describe a clean and simple method for integrating \tcltk\ and \CXX\ 
in order to provide the means to create graphical user interfaces for
Orbix-based applications, both clients and servers.  In fact, the method
is a general one which allows the creation of a \tcltk\ GUI for any \CXX\ 
program.

The method allows \CXX\ functions to be invoked from a \tcltk\ script and
allows \tcltk\ commands to be scheduled for evaluation from \CXX.  
Although most \omniParTcl\ applications will be multithreaded,
all \tcltk\ processing is guaranteed to be confined to a single thread.
No modification to the Tk scheduler is required.
\end{abstract}

\section{Introduction}

\omniParTcl\ (pronounced ``omni-{\em particle}'' and intended to
evoke the idea of ``{\em Par\/}allel {\em \tcl\/}'') provides a mechanism
for attaching a graphical front-end written in \tcltk\ to a \CXX\ 
program.  It was motivated by the perceived need for
better-than-command-line interfaces for Omni/Orbix applications.  In
discussion, the following criteria were established for any integration
of \tcltk\ and \CXX:

\begin{itemize}

\item 
It must be possible for the \tcltk\ interface to run in parallel with
\CXX\ computation.  This requirement stems from potential delays in
Orbix RPCs, but it is also desirable in the general case to allow
the user to continue to interact with an application during
lengthy computations of any sort.

\item
All \tcltk\ primitives should be invoked from a single thread.  This
frees us---at least to a certain extent---from worries about
thread-aware i/o and, more importantly, prevents dangerous inter-thread
recursion which could otherwise easily occur given the diverse range of
execution triggers in \tcltk.

\item
Polling should not be used in synchronizing \tcltk\ and \CXX\ 
threads.  This obviates the need to predetermine the balance between
interactive response time and CPU wastage on polling.  More generally,
it also removes the danger of multiple \omniParTcl\ applications
dominating the CPU simply due to polling.

\item 
If possible, no modifications to the Tk scheduler should be required.
That is, we would like our \tcltk\ interface to be managed by a {\tt
wish} which has been constructed using only the standard
hooks.  This makes it much more likely that the integration technique
will work with future releases of \tcltk, and also means that upgrading
to such future releases will involve little, if any, work.

\end{itemize}

\noindent
\omniParTcl\ has been designed to meet all of these criteria.

\section{Using \omniParTcl}

To write an application using \omniParTcl, you create two related pieces
of code:

\begin{enumerate}

\item The application code.  This is the \CXX\ source of your
application, which will consist of any number of \CXX\ functions and
may, of course, include Orbix code.  Your application code must define
{\tt Tcl\_Appinit()} and {\tt main()}.  Your {\tt Tcl\_Appinit()} must
call {\tt omniParTcl\_Init()} which takes the same form as {\tt
Tcl\_Init()} and {\tt Tk\_Init()}.  Your {\tt main()} must call
{\tt Tk\_Main()}.

By using {\tt Tcl\_CreateCommand} within your definition of {\tt
Tcl\_Appinit}, you will nominate certain of your \CXX\ functions as
callable from \tcl.  Usually these \tcl-callable \CXX\ functions will
launch one or more parallel threads and return control to the main
\tcltk\ thread, although simple functions can be run in the main thread
if their execution does not interfere unduly with the \tcltk\ user
interface.

For communication in the other direction, \omniParTcl\ provides two
\CXX\ functions which schedule \tcltk\ commands for evaluation in the
primary thread.  

\begin{itemize}
\item
{\tt omniTclMeAndRun(s)} schedules \tcltk\ script {\tt s}
for evaluation as soon as possible and returns control
immediately to the calling \CXX\ thread.  
\item
{\tt omniTclMeAndWait(s)}
schedules the script for evaluation and does not return control to the
caller until the script has been evaluated.  
\end{itemize}

\noindent
Clearly, {\tt omniTclMeAndWait()} cannot be used by a
\CXX\ function running in the main thread.

\item

The \tcltk\ script.  This implements your graphical user interface, and will
contain calls to the \CXX\ functions nominated in {\tt Tcl\_Appinit()} as
callable from \tcltk.  One way of running your program is with {\tt myApp -file
<script>} where {\tt myApp} is the name of your compiled \CXX\ application code
and {\tt <script>} is the name of your \tcltk\ script.  Alternatively you can
use the {\tt TclScriptExecutable} rule in your {\tt dir.mk}.  This will
generate a new script with an appropriate \verb|#!| line at the top.

\end{enumerate}

\section{Building \tcltk\ scripts in \CXX}

To make it easier to build complex \tcltk\ scripts for scheduling
with the {\tt TclMe} functions, \omniParTcl\ provides a \CXX\ class
wrapper for the {\tt Tcl\_DString} routines.  In fact, the wrapper also
extends {\tt Tcl\_DString} functionality by providing overloaded
members for appending longs and doubles, and by extending {\tt
appendElement} to behave exactly like {\tt append}.

Thus, 
you can use {\tt append()} and {\tt appendElement()} to
attach strings, longs, and doubles to a growing DString.  E.g.

\begin{verbatim}
     tclDString ds;
     double d;
     long i;

     ...

     ds.append("set mylist ");
     ds.startSublist();
     ds.appendElement(i);
     ds.appendElement(d);
     ds.endSublist();

     omniTclMeAndRun(ds.value());
\end{verbatim}

Note also that there are appropriate default arguments for certain
member functions so that, for example, the following equivalences
hold,

\begin{verbatim}
     ds.append("xyz")     <==>     ds.append("xyz", -1)

     ds.append(i)         <==>     ds.append(i,"%d")  

     ds.append(d)         <==>     ds.append(d,"%g")
\end{verbatim}

\noindent
and similar ones hold for the {\tt appendElement()} versions.  Thus, by
default you append an entire string, a long in decimal format, and a
double in the smallest accurate way, but by providing the second
argument you can append only part of string, longs in hex or octal, and
doubles in any format you choose.

See the \tcltk\ manual for more detail about the {\tt Tcl\_DString} functions.

\section{How it works}

\omniParTcl\ really just organizes a few simple conventions.   
First, {\tt omniParTcl\_Init} sets things up by
performing the following tasks:

\begin{itemize}
\item
It creates a pipe and uses the function {\tt Tk\_CreateFileHandler} to
arrange to call the function {\tt cxxTclScriptHandler()} whenever data
is available to read from the pipe.

\item
It creates two mutexes, one for protecting access to the pipe and 
another for protecting access to a queue of scripts awaiting evaluation.
\end{itemize}

\noindent
The essential behaviour of \omniParTcl\ is then determined by the
interplay between the two script-evaluation functions,
{\tt omniTclMeAndRun()} and {\tt omniTclMeAndWait()}, and 
the function {\tt cxxTclScriptHandler()}.  Recall that the latter
is always invoked within the primary \tcltk\ thread, while the
two {\tt TclMe} functions will typically be invoked in
other threads.

\begin{itemize}

\item Each of the two {\tt TclMe} functions puts the supplied
script onto the same FIFO queue of scripts awaiting evaluation, and
writes a byte to the pipe.  {\tt omniTclMeAndRun()} simply returns at
this point, while {\tt omniTclMeAndWait()} creates a condition variable,
puts it on the queue with the script, and then waits on it.

\item {\tt cxxTclScriptHandler()} is invoked within the primary
\tcltk\ thread when there is something to read from the pipe.  It
performs the following steps until the queue is empty:

\begin{itemize}

\item It consumes a byte from the pipe.

\item It removes a script from the queue and
evaluates it using {\tt Tcl\_GlobalEval}.

\item If the script is paired with a condition variable, 
it signals on the condition.
\end{itemize}

\end{itemize}

\noindent
The interplay amongst these functions (in particular, their access
to the pipe and script queue) is, of course, properly mutex-protected.

Er, that's it.

\section{Non-graphical and mixed applications}

Because the \omniParTcl\ developer writes his or her own {\tt main()},
the application's behaviour can be tailored in various ways.  In
particular, by calling {\tt Tcl\_Main} instead of {\tt Tk\_Main}
and by omitting the call to {\tt Tk\_Init} in 
{\tt Tcl\_Appinit()}, the developer can write an application which
uses \tcl\ without \tk\ to create the (presumably non-graphical) interface.

Perhaps more usefully, the developer could provide both types of interface
and allow a command-line flag to dictate which one is used.  

\section{Where is it now?}

\omniParTcl\ is installed in the Omni Development Environment.
The source can be found in:

\begin{verbatim}
/project/omni/versionX.Y/include/omniParTcl.h
/project/omni/versionX.Y/src/lib/omniParTcl/*
\end{verbatim}

In order to encourage migration to the new \tcltk\ libraries,
\omniParTcl\ is built for \tcl 7.4 and \tk 4.0 and will not
work with earlier versions.  In fact, the Makefile variables 
\verb|OMNIPARTCL_CPPFLAGS| and \verb|OMNIPARTCL_LIB| automatically provide
include files and libraries for \tcl 7.4 and \tk 4.0.  However, if there is a
pressing need to use \omniParTcl\ with \tcl 7.3 and \tk 3.6, it can
easily be built for these.

Note that the {\tt omnithread} library is used for all \omniParTcl\ 
applications, even those which have only a single thread.

\section{An Example}

The code for a simple \omniParTcl\ application appears on the following
pages.  The source consists of three files.

\begin{enumerate}
\item {\tt pttwish.cc}: the \CXX\ application code.
\item {\tt ptt.tcl}: the \tcltk\ interface script.
\item {\tt dir.mk}: the makefile to build the example in the Omni
                       Development Environment.
\end{enumerate}

\noindent
These files can be found in
{\tt /project/omni/versionX.Y/src/tests/partcl}.

\medskip

In order to emphasize the \omniParTcl\ side of things, this example does
not use Orbix.  However, no extra magic is required to incorporate Orbix
code into an \omniParTcl\ application, and examples which use Orbix
and which have much more complex user interfaces can be found by asking the
author of this document.

\newpage
\begin{verbatim}
//---------------------------------------------------------------------------
// pttwish.cc - Simple test of omniParTcl. We just repeatedly update a couple
//          of variables on the C++ side and monitor their values in Tcl/Tk 
//          labels.  To test communication the other way, we allow the user 
//          to change the output format for one of the variables.
//---------------------------------------------------------------------------

#include <string.h>
#include "omniParTcl.h"

static omni_thread *threadOne;
static omni_thread *threadTwo;

static char *fmt;
static omni_mutex fmtMutex;

//---------------------------------------------------------------------------

void threadOneFunc(void *arg)
{
  tclDString ds;
  static long i = 0;
  
  while (1) { // Repeatedly do some Tcl.
      
    ds.init();
    ds.append("set tclVarOne ");

    //
    // By using appendElement() rather than append() we ensure that
    // any format string (e.g. one containing whitespace) will work.
    //
    fmtMutex.acquire();
    ds.appendElement(++i,fmt);      
    fmtMutex.release();    
    
    omniTclMeAndRun(ds.value());
    ds.free();
      
    omniTclMeAndWait("update");
    
    omni_thread::sleep(0,100000000); // Sleep a bit
  }
}
\end{verbatim}
\newpage
\begin{verbatim}
//---------------------------------------------------------------------------

void threadTwoFunc(void *arg)
{
  tclDString ds;
  static double d = 0.999;
  
  while (1) { // Repeatedly do some Tcl.

    ds.init();
    ds.append("set tclVarTwo ");
    ds.append(d);
    
    omniTclMeAndRun(ds.value());
    ds.free();

    d -= 0.001;
    
    omni_thread::sleep(1,0); // Sleep a second
  }
}

//---------------------------------------------------------------------------

int startThreadOne(ClientData clientData, Tcl_Interp *interp,
                   int argc, char *argv[])
{
  if ((threadOne = omni_thread::create(threadOneFunc,NULL)) == NULL) {
    Tcl_SetResult(interp, "ptt: cannot create threadOne", TCL_VOLATILE);
    return TCL_ERROR;
  }
  return TCL_OK;
}

//---------------------------------------------------------------------------

int startThreadTwo(ClientData clientData, Tcl_Interp *interp,
                   int argc, char *argv[])
{
  if ((threadTwo = omni_thread::create(threadTwoFunc,NULL)) == NULL) {
    Tcl_SetResult(interp, "ptt: cannot create threadTwo", TCL_VOLATILE);
    return TCL_ERROR;
  }
  return TCL_OK;
}
\end{verbatim}
\newpage
\begin{verbatim}
//---------------------------------------------------------------------------

int setFormat(ClientData clientData, Tcl_Interp *interp,
              int argc, char *argv[])
{
  //
  // Make sure we've been passed a new format string.  
  // If so, use it to update the format.  Otherwise,
  // leave it as-is.
  //
  
  if (argc >= 2) {
    delete fmt;
    fmtMutex.acquire();
    fmt = new char[strlen(argv[1])+1];
    strcpy(fmt,argv[1]);
    fmtMutex.release();
  }
  
  return TCL_OK;
}

//---------------------------------------------------------------------------

int Tcl_AppInit(Tcl_Interp *interp)
{
  //
  // For omniParTcl applications, Tcl_Appinit must call omniParTcl_Init()
  // as well as the Tcl and Tk init routines.   Any C++ functions which
  // are to be callable from Tcl must be specified in a call to
  // Tcl_CreateCommand().
  //

  Tk_Window main;
  main = Tk_MainWindow(interp);

  if (Tcl_Init(interp) == TCL_ERROR) {
    return TCL_ERROR;
  }
  if (Tk_Init(interp) == TCL_ERROR) {
    return TCL_ERROR;
  }
  if (omniParTcl_Init(interp) == TCL_ERROR) {
    return TCL_ERROR;
  }

  // tcl_RcFileName = ".mytclrc"; // No RC file for this app.

  //
  // Make our C++ functions available in Tck/Tk.
  //
  Tcl_CreateCommand(interp, "CXXstartThreadOne", startThreadOne,
                    (ClientData) NULL, (Tcl_CmdDeleteProc *) NULL);

  Tcl_CreateCommand(interp, "CXXstartThreadTwo", startThreadTwo,
                    (ClientData) NULL, (Tcl_CmdDeleteProc *) NULL);

  Tcl_CreateCommand(interp, "CXXsetformat", setFormat,
                    (ClientData) NULL, (Tcl_CmdDeleteProc *) NULL);

  //
  // Application specific initialization.  Default format for
  // the first variable is standard decimal.
  //
  fmt = new char[3];
  strcpy(fmt,"%d");
  
  return TCL_OK;
}

//---------------------------------------------------------------------------

int main(int argc, char *argv[])
{
  //
  // omniParTcl is designed for use with Tcl7.4/Tk4.0 which means that
  // the user must write his or her own main().  This provides extra
  // flexibility.  For example, a command-line switch could determine
  // whether the application used a graphical interface (in which case
  // it would call Tk_Main) or an ascii interface (in which case it
  // would call Tcl_Main.)
  //
  // For typical Tk-based graphical interfaces, main() simply calls
  // Tk_Main() and returns 0 as below.  In this case, the program
  // behaves like a standard wish and would typically be invoked
  // using the "# <program> -f" construct at the top of a Tcl/Tk
  // source file.
  //
  
  Tk_Main(argc, argv, Tcl_AppInit);  // Run the Tk scheduler.
  return 0;	                     // Make the compiler happy.
}
\end{verbatim}


\newpage
\begin{verbatim}
#
# Simple test of omniParTcl.  We just repeatedly update a couple of 
# variables on the C++ side and monitor their values in Tcl/Tk labels.
# To test communication the other way, we allow the user to change the
# output format for one of the variables.
#

set tclVarOne 0
set tclVarTwo 0

CXXstartThreadOne
CXXstartThreadTwo

label .l1 -textvariable tclVarOne -bg cyan
label .l2 -textvariable tclVarTwo -bg green

button .q -text "Quit" -command exit
button .b -text "Change format" -command changeFormat

pack .q .l1 .l2 .b -fill both
update

proc changeFormat {} {
  pack forget .b  
  label .fl -text "New format string:"
  entry .fe 
  pack .fl .fe -side left
  update
  bind .fe <Return> {CXXsetformat [.fe get] \
                     destroy .fe; destroy .fl \
                     pack .b; update}
}
\end{verbatim}


\newpage
\begin{verbatim}
#
# dir.mk for simple omniParTcl test.
#

CXXSRCS = pttwish.cc

DIR_CPPFLAGS = $(OMNIPARTCL_CPPFLAGS)

script = $(patsubst %,$(TclScriptPattern),ptt)

wish = $(patsubst %,$(BinPattern),pttwish)

all:: $(wish) $(script)

$(wish): pttwish.o $(OMNIPARTCL_LIB_DEPEND)
        @(libs="$(OMNIPARTCL_LIB)"; $(CXXExecutable))

$(script): ptt.tcl
        @(wish="$(wish)"; $(TclScriptExecutable))

export: $(wish) $(script)
        @$(ExportExecutable)

clean::
        $(RM) $(wish) $(script)
\end{verbatim}


\end{document}


